\Large \bf Module-level Dynamic Analysis

### 2019-12-19 paper/draft.tex
Add implementation details

\section{Implementation}
\label{impl}

>>>>

\section{Implementation: Lya}
\label{impl}

We implemented \sys for version 
We show where other implementations would diverge in table 1

The following subsections outline how module systems handle modules at runtime, by exemplifying the internals of Node.js.
It sketches the Node.js runtime~\sx{a}, the use of the module system~\sx{b}, and the internals of how module loading works~\sx{c}.

(1) why node.js
(2) Create a table of compatibility and show individual aspects

\heading{Module System}
Node.js' module system is implemented entirely in JavaScript.
It exposes \ttt{require}, a global-looking function for importing modules.
\fixme{Figure} above demonstrates the use of \ttt{require}.

% \begin{lstlisting}[language=js,mathescape,upquote=true]
% // -------------- [./main.js] -------------
% // importing point
% let Point = require("./point.js");
% Point.create(1, 1).print(); // => [1, 1]
% 
% // -------------- [./point.js] ------------
% let Point = function Point (x, y) {
%     this.x = x; this.y = y;
% };
% Point.prototype.print = function print () {
%   console.log([this.x, this.y]);
% };
% module.exports = {
%     create:  function create (x, y) {
%         return new Point(x, y);
%     }
% };
% \end{lstlisting}

In the example above, the main module (\ttt{main.js}) imports the \ttt{point.js} module using the \ttt{require} statement (line 3)
Functionality from the exporting module (\ttt{point.js}) that is expected to become available to the importing module (\ttt{main.js}) is assigned to a special \ttt{module.exports} object (line 13);
  the rest is module-private functionality.
Files and modules are in one-to-one correspondence (each file is treated as a separate module).
Method \ttt{require} is synchronous (\ie blocking):
  it will block execution until the module specified is loaded.
% ---and can be used to load built-in modules such as \ttt{http} and \ttt{fs}.
The module system is implemented in the \ttt{module} built-in module~\sx{c}, which locates, wraps, compiles, and executes the specified file.

\sys hooks into \ttt{module} and alters the return value of the \ttt{require} call that imports \ttt{point.js} (line 3).

\subsection{Implementation of the Module System}
\label{c}

At a very high level, loading a fresh module with \ttt{require("foo");} corresponds to the following five stages:

\begin{enumerate}
% \def\labelenumi{\arabic{enumi}.}
\item Resolution: identify the file to which the module specified corresponds, and locate it in the filesystem.
\item Loading: depending on the file type, use the corresponding loader (\eg V8 compiler for \ttt{js}, \ttt{JSON.parse} for \ttt{json} \etc).
\item Wrapping: wrap the module so that module-globals get encapsulated and Node.js globals (\eg \ttt{require}) get resolved.
\item Evaluation: evaluate the wrapped module in the current context, so that global names and top-level objects get resolved correctly.
\item Caching: add the module to a handful of module-related cache structures, for purposes of consistency and performance.
\end{enumerate}

% The resolution algorithm is somewhat convoluted, because it depends on several different facts (including the type of the file requested, whether it is a globally installed, whether it is a directory \etc).
% It does not require any \sys support beyond copying the directory that contains the source code of the all the modules onto the remote host.
\sys interposes on all of these steps to facilitate transformations.
Wrapping (3) and evaluation (4) are particularly interesting, because they allow \sys to interpose at the module boundary during runtime.
Before a module's code is evaluated, the Node.js module loader wraps the module so that
  (i) it keeps top-level variables (defined with \ttt{var}, \ttt{const} or \ttt{let}) scoped to the module rather than the global object; and
  (ii) it provides some global-looking variables that are actually specific to the module, such as the \ttt{module} and \ttt{exports} objects that the implementor can use to export values from the module and convenience variables---such as \ttt{\_\_filename} and \ttt{\_\_dirname}  containing the module's absolute filename and directory path, respectively.
True globals remaining are
  (i) the global objects as defined by the EcmaScript standard (\eg \ttt{Object}, \ttt{Function}, \ttt{Math}); and
  (ii) Node.js-specific globals (\eg \ttt{console}, \ttt{process}, \ttt{timer}).
These globals require further interposition.

% \begin{lstlisting}[language=js,mathescape,upquote=true]
% // Node.js will wrap a module with a function,
% // so as to bring certain names into scope
% // before compiling/evaluating code.
% let wrapped = "function (" +
%         "exports, require, module, " +
%         "__filename, __dirname, CTX) {" +
%     "let Math = CTX.Math" +
%     "let console = CTX.console" +
%     //...[more definitions]
%     moduleSource +
%   "});"
% \end{lstlisting}

\sys hooks into the wrapper function (the last variable in the function definition, \ttt{CTX}).
This trivial source-to-source transformation re-defines global variables as module-locals and initializes them with \sys-augmented values.
For example, \ttt{console} in the context of the module will be an \sys-created object that allows \sys to interpose on it.
Evaluation of the module passes an additional value to this function, which is the modified context.
As a result, any changes to the top-level objects and any global variables are accessible from the within the module.

% \begin{lstlisting}[language=js,mathescape,upquote=true]
% //Input: module ID e.g., absolute filepath
% let load = function (ID) {
%   if (cache[ID]) {
%      return cache[ID];
%   }
%   let m = {
%     exports: {}, id: ID, dir: path.resolve(ID)
%   };
%   let cm = v8.compile(wrapped);
%   let ii = iris.getImplicit(ID);
%   let c = iris.freshContext(ii);
%   cm(m.exports, this.require, m, ID, m.dir, c);
%   let ei = iris.getExplicit(ID);
%   m.exports = iris.wrap(m.exports, [ei] );
%   cache[ID] = m;
%   return m.exports;
% }
% \end{lstlisting}

The \ttt{load} method in the \ttt{Module} module combines evaluation (line 10) and caching (line 14) of the wrapped module.
After evaluation, invoking the compiled function generates the value that is assigned to \ttt{module.exports} from within the module (line 12).
\sys passes a freshly constructed context at that invocation, modified according to the implicit segment of the (PIC??) corresponding to the module being loaded.
Before returning the value of \ttt{module.exports}, \sys transforms it according to the explicit segment of the (PIC??) corresponding to the module being loaded.
Finally, the results of the entire process are placed into the module cache for later use.

\
